---
output: github_document
---

# Integrating H2O & tidyverse

During [LatinR2019](https://latin-r.com/), several people were asking Erin Ledell about tidyverse and H2O integration, she said that it was actually a work in mind. So I thought, if my community would enjoy it, maybe as a community we should do it -that's what open source is about-.

By installing the `tidyh2o` package:

```{r eval = FALSE}
if (!require("remotes")) {
  install.packages("remotes")
}
remotes::install_github("jcrodriguez1989/tidyh2o")
```

we can use `h2o` in a tidyverse style:

```{r echo = FALSE}
unloadNamespace("dplyr")
unloadNamespace("tidyh2o")
```

```{r}
library("dplyr")
library("h2o")
library("tidyh2o")

h2o.init()
h2o.no_progress() # Turn off progress bars for notebook readability

# Use data file from GitHub
data_file <- 
  "https://raw.githubusercontent.com/ledell/LatinR-2019-h2o-tutorial/master/data/loan.csv"

data <- h2o.importFile(data_file) # 163,987 rows x 15 columns
data %>% dim()

# Optional (to speed up the examples)
data %<>% sample_n(30000)

data %<>%
  mutate(bad_loan = as.factor(bad_loan)) #encode the binary repsonse as a factor
data$bad_loan %>% h2o.levels() #optional: this shows the factor levels

data %>% h2o.describe()

splits <- data %>% h2o.splitFrame(
  ratios = c(0.7, 0.15), # partition data into 70%, 15%, 15% chunks
  destination_frames = c("train", "valid", "test"), # frame ID (not required)
  seed = 1) # setting a seed will guarantee reproducibility
train <- splits[[1]]
valid <- splits[[2]]
test <- splits[[3]]

train %>% nrow
valid %>% nrow
test %>% nrow

y <- "bad_loan"
#remove the interest rate column because it's correlated with the outcome
x <- setdiff(names(data), c(y, "int_rate"))
print(x)

aml <- h2o.automl(y = y, x = x,
                  training_frame = train,
                  max_models = 10,
                  seed = 1)
(lb <- aml@leaderboard)
print(lb, n = nrow(lb))

aml_perf <- aml@leader %>% h2o.performance(newdata = test)
aml_perf %>% h2o.auc()

preds_fm <- aml %>% predict(newdata = test)
preds_fm %>% head()
preds_ld <- aml@leader %>% predict(newdata = test)
preds_ld %>% head()

table(as.vector(preds_fm$predict), as.vector(preds_ld$predict))

# more `dplyr` stuff we can do
data %>%
  select("annual_inc", "loan_amnt") %T>%
  plot() %>%
  colSums(na.rm = TRUE)

data %<>%
  filter(term == "36 months") %>%
  mutate(high_loan = loan_amnt > median(loan_amnt)) %>%
  arrange(desc(loan_amnt))

data %>%
  group_by(home_ownership) %>%
  summarise(
    n = n(),
    loan_amnt = mean(loan_amnt, na.rm = TRUE)
  ) %>%
  arrange(desc(loan_amnt))

h2o.shutdown(prompt = FALSE)
```

## How is this implementation working?

To integrate `h2o` and the `tidyverse`, I thought that as a very first attempt, we could redefine `magrittr` `pipe` operators. The idea would be: 

* when calling a pipe with a non-`H2OFrame`, just use the regular `magrittr` pipes.

* when calling a pipe with an `H2OFrame`, then convert it to `data.frame`, apply all operators, and if the result is a `matrix`/`data.frame` then reconvert it to an `H2OFrame`.

These code editions can be seen at [https://www.diffchecker.com/LyjKierQ](https://www.diffchecker.com/LyjKierQ).

### Pros:

* It took a few lines of code to implement `tidyh2o`.

### Cons:

* This is really inefficient because, for each pipe operator, it is doing one or two `H2OFrame`~`data.frame` conversions.

* It is redefining the so used `pipe` operators!! And as a side effect, it depends on the order that libraries are loaded (`tidyh2o` must be loaded after `magrittr`, `dplyr`, etc.).

## Optimal solution

I think the best solution would be to redefine the `H2OFrame` class, make it be a `data.frame`/`tibble`, and keep all the Java-H2OFrame information as attributes. In this way, all `tidyverse` functions will work as usual. And then, just update the Java-H2OFrame when needed to do some ML calculations (at Java workspace).

This, would be a big, and very nice collaborative work. Most of the `h2o` functions should be -easily- redefined.
